Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2019-03-20 mpaper/paper.tex
Introduction!

\begin{abstract}
  We prove the validity of performing variational inference on the Gaussian
  process-based inverse reinforcement learning model that preserves and
  approximates the full joint probability distribution of rewards across states
  of the Markov decision process.
\end{abstract}

\section{Introduction} % first page
% TODO: rewrite and make it good
% 1. Problem (with examples, intuitive)
% 2. Its importance/relevance/motivation
% 3. My idea
% 4. List of contributions
% Forward-reference which section deals with what

Inverse reinforcement learning (IRL)---a problem proposed by Russell in
1998 \cite{DBLP:conf/colt/Russell98}---asks us to find a reward function for a
Markov decision process (MDP) that best explains a set of given demonstrations.
IRL is important because reward functions can be hard to define manually
\cite{DBLP:conf/icml/PieterN04,DBLP:journals/corr/abs-1806-06877}, and rewards
are not entirely specific to a given environment, allowing one to reuse the same
reward structure in previously unseen environments
\cite{DBLP:journals/corr/abs-1806-06877,DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}.
Moreover, IRL has seen a wide array of applications in autonomous vehicle
control \cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16} and
learning to predict another agent's behaviour
\cite{DBLP:journals/ai/BogertD18,DBLP:conf/aaai/VogelRGR12,ziebart2008maximum,DBLP:conf/huc/ZiebartMDB08,DBLP:conf/iros/ZiebartRGMPBHDS09}.
Most approaches in the literature make a convenient yet unjustified assumption
that the reward function can be expressed as a linear combination of features.
One proven way to abandon this assumption is by representing the reward function
as a Gaussian process (GP)
\cite{DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11,DBLP:journals/corr/abs-1208-2112}.
The original approach used maximum likelihood estimation
\cite{DBLP:conf/nips/LevinePK11}, whereas we use variational inference (VI)
instead, which learns approximate posterior probability distributions instead of
point estimates. This approach can prove useful in three major ways:
\begin{enumerate}
\item Modelling full posterior distributions for various parameters can result
  in more precise reward predictions, as the model simply holds more
  information.
\item Having variance estimates for rewards can direct our choice in what data
  should be collected next.
\item An approximate Bayesian treatment of many parameters in the model guards
  against overfitting \cite{DBLP:conf/uai/JinDAS17}.
\end{enumerate}

@@

\section{The Model}

@@

\section{Theoretical Justification}

@@

The typical way to optimise a quantity (the ELBO, in this case) involves
computing its gradient. Unfortunately, the term $\mathbb{E}[v]$ in $\mathcal{L}$
complicates the situation. The goal of this section is to show how Theorem
\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
anyway\footnote{This technique is inspired by black box VI
  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
  problem and requires significantly more work to prove correctness.}. After
showing that the theorem applies to our situation, we can express
$\nabla\mathbb{E}[v]$ as
\begin{align*}
  \nabla\mathbb{E}[v] &= \nabla \iint q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
                      &= \iint \nabla[v q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})]\dx \\
                      &= \iint \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})} q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
                      &\approx \frac{1}{S} \sum_{s=1}^S \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})},
\end{align*}
which can be computed by drawing $S$ Monte Carlo samples $(\mathbf{u}_s,
\mathbf{r}_s) \sim \approximation$.

@@

\section{Evaluation} % showing deep understanding

@@

@@

Table ?? further summarises some of the recent and/or
influential GP approximation approaches that might be relevant to our situation.
For many of the variables in our model (including the inducing points and the
hyperparameters of the covariance function), we have three possible ways of
handling them:
\begin{itemize}
\item Provide a full Bayesian treatment by defining a prior probability
  distribution. While this option can easily prevent overfitting, the choice of
  prior can be difficult to justify.
\item Fix the value of the variable. We can then optimise the value outside of
  the main algorithm, e.g., by running the algorithm with a set of possible
  values and keeping the value that produces the best performance. This option
  can be slow if we are interested in finding the optimal value, but could be
  appropriate if the variable has little impact on the algorithm's performance.
\item Treat the variable as a variational parameter. This method will
  efficiently optimise the value of the variable, but is vulnerable to
  overfitting. Furthermore, this requires us to find the derivative of the ELBO
  w.r.t. the variable, which could be difficult.
\end{itemize}

@@

@@

\subsection{Further Work}

>>>>

\begin{abstract}
  We prove the validity of performing variational inference on the Gaussian
  process-based inverse reinforcement learning model that preserves and
  approximates the full joint probability distribution of rewards across states
  of the Markov decision process.%TODO: finish this.
  According to Simon Peyton Jones, an abstract should address
  four key questions. First, what is the problem that this
  paper tackles? Second, why is this an interesting problem?
  Third, what is the solution this paper proposes?
  Finally, why is the proposed solution a good one?
\end{abstract}

\section{Introduction} % first page
% 1. Problem (with examples, intuitive)
% 2. Its importance/relevance/motivation
% 3. My idea
% 4. List of contributions
% Forward-reference which section deals with what

Imagine using a machine learning (ML) algorithm to teach a robot how to move
around people so that it learns to predict where people are going and adjust its
path accordingly. The ML algorithm would use data about various possible
situations. But do we have enough data to ensure reasonably optimal behaviour?
Perhaps the robot behaves well in most situations, but fails in less common
scenarios. Can the ML model itself describe its weaknesses so that we could
ensure it is exposed to sufficiently many uncommon or difficult situations?

This learning problem
\cite{DBLP:journals/ijsr/KimP16,DBLP:journals/ijrr/KretzschmarSSB16} as well as
many others have benefited from an approach called \emph{inverse reinforcement
  learning} (IRL) (also known as inverse optimal control). IRL proposes a way to
learn behaviour from \emph{demonstrations} that typically come from human
actions. More formally, the IRL problem asks us to find a reward function for a
Markov decision process (MDP), where demonstrations are encoded as sets of
state-action pairs.

IRL is an important problem because adjusting the reward function by hand is
often unwieldy, since human behaviour often depends on many factors in
complicated ways
\cite{DBLP:conf/icml/PieterN04,DBLP:journals/corr/abs-1806-06877}. Moreover,
learning the reward function rather than the policy itself makes the model more
transferable to new environments---a minor change in the environment can
reorganise the whole policy but only have a local effect in the reward structure
\cite{DBLP:journals/corr/abs-1806-06877,DBLP:conf/uai/JinDAS17,DBLP:conf/nips/LevinePK11}.
IRL has been used to teach helicopters how to perform tricks
\cite{DBLP:conf/nips/AbbeelCQN06}, predict taxi destinations
\cite{DBLP:conf/huc/ZiebartMDB08}, and make driving safer and more efficient by
predicting pedestrian movement \cite{DBLP:conf/iros/ZiebartRGMPBHDS09} and the
driver's intentions \cite{DBLP:conf/aaai/VogelRGR12}.

% TODO: need to rewrite the first sentence
However, most IRL models in the literature make a convenient yet unjustified
assumption that the reward function can be expressed as a linear combination of
features
\cite{DBLP:conf/icml/PieterN04,DBLP:conf/uai/NeuS07,DBLP:conf/icml/NgR00,ziebart2008maximum}.
This assumption makes the models unable to represent many reward structures. Out
of the non-linear models proposed to date, none can answer the questions posed
in the first paragraph. Quite often, the models assume that rewards have no
variance \cite{DBLP:conf/nips/LevinePK11,DBLP:conf/uai/JinDAS17}. In this paper,
we show how that assumption can be lifted by switching from maximum likelihood
estimation to \emph{variational inference} (VI), i.e., we approximate the
posterior distribution of the model by optimising the parameters of a simpler
distribution to make it similar to the posterior. This approach can prove useful
in four major ways:
\begin{enumerate}
\item By working with full distributions instead of point estimates, we can
  expect more precise reward predictions.
\item Variance estimates can be used to guide what data should be collected
  next, i.e., if the rewards of some states have abnormally high variance, we
  might want to expose the model to more data visiting those and surrounding
  states.
\item Variances estimates can also be used to judge whether we can trust the
  predictions of the model or, perhaps, the model could benefit from some
  adjustments or more data.
\item By adopting a more Bayesian approach, we automatically incorporate Occam's
  razor into the model that guards against overfitting
  \cite{DBLP:conf/uai/JinDAS17}.
\end{enumerate}

Our main contribution is a lengthy proof in Section \ref{sec:proof} that shows
how VI can be applied to the maximum-entropy IRL model with Gaussian processes
(GPs) proposed by Levine et al. \cite{DBLP:conf/nips/LevinePK11}. We describe
how we adapt the model and set up the VI problem in Section \ref{sec:model}.
Finally, in Section \ref{sec:evaluation} we examine the convergence properties
of our model and its ability to deduce optimal policies in practice.

@@

\section{The Model} \label{sec:model}

@@

\section{Theoretical Justification} \label{sec:proof}

@@

The typical way to optimise a quantity (the ELBO, in this case) involves
computing its gradient. Unfortunately, the term $\mathbb{E}[v]$ in $\mathcal{L}$
complicates the situation. The goal of this section is to show how Theorem
\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
anyway\footnote{This technique is inspired by black box VI
  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
  problem and requires significantly more work to prove correctness.}. After
showing that the theorem applies to our situation, we can estimate
$\nabla\mathbb{E}[v]$ with
\begin{align*}
  \nabla\mathbb{E}[v] &= \nabla \iint q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
                      &= \iint \nabla[v q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})]\dx \\
                      &= \iint \frac{\nabla[v q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})]}{q(\mathbf{r} \mid \mathbf{u})q(\mathbf{u})} q(\mathbf{r} \mid \mathbf{u}) q(\mathbf{u})\dx \\
                      &\approx \frac{1}{S} \sum_{s=1}^S \frac{\nabla[v q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)]}{q(\mathbf{r}_s \mid \mathbf{u}_s)q(\mathbf{u}_s)},
\end{align*}
which can be computed by drawing $S$ samples $\{(\mathbf{u}_s,
\mathbf{r}_s)\}_{s=1}^S$ from $\approximation$.

@@

\section{Evaluation} \label{sec:evaluation}
% showing deep understanding

@@

% apprenticeship learning, imitation learning
%a problem proposed by Russell in 1998 \cite{DBLP:conf/colt/Russell98}

@@

For many of the variables in our model (including the inducing points and the
hyperparameters of the covariance function), we have three possible ways of
handling them:
\begin{itemize}
\item Provide a full Bayesian treatment by defining a prior probability
  distribution. While this option can easily prevent overfitting, the choice of
  prior can be difficult to justify.
\item Fix the value of the variable. We can then optimise the value outside of
  the main algorithm, e.g., by running the algorithm with a set of possible
  values and keeping the value that produces the best performance. This option
  can be slow if we are interested in finding the optimal value, but could be
  appropriate if the variable has little impact on the algorithm's performance.
\item Treat the variable as a variational parameter. This method will
  efficiently optimise the value of the variable, but is vulnerable to
  overfitting. Furthermore, this requires us to find the derivative of the ELBO
  w.r.t. the variable, which could be difficult.
\end{itemize}

@@

% Show how we achieved our claimed contributions

@@

\subsection{Further Work}
% How much does a reward have to change to change the policy? Could be an
% interesting use case for variance predictions.
